On the presentation of orthogonal groups over polynomial rings

### 2019-05-02 dl-paper.tex
More coherent notation, added couple of easy formulas

@@

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of type 3.
\begin{comment}
Let us draw the matrix for $z_\alpha(m, X\xi)$:
\[ \left(\begin{array}{cc} 1 & 0 \\ -X\xi & 1 \end{array} \right) \cdot \left(\begin{array}{cc} 1 & m \\ 0 & 1 \end{array} \right) \cdot \left(\begin{array}{cc} 1 & 0 \\ X\xi & 1 \end{array} \right) = 
 \left(\begin{array}{cc} 1 & m \\ -X\xi & 1 - mX\xi \end{array} \right) \cdot \left(\begin{array}{cc} 1 & 0 \\ X\xi & 1 \end{array} \right) = \left(\begin{array}{cc} 1 + mX\xi & m \\ -mX^2\xi^2 & 1 - mX\xi \end{array} \right) \]
Now let us draw the matrix for $c_\alpha(m, \xi)$:
\[ \left(\begin{array}{cc} 1 & m \\ 0 & 1 \end{array} \right) \left(\begin{array}{cc} 1 + mX\xi & -m \\ mX^2\xi^2 & 1 - mX\xi \end{array} \right) = 
 \left(\begin{array}{cc} 1 + mX\xi + m^2X^2\xi^2 & -m^2X\xi \\ mX^2\xi^2 & 1 - mX\xi \end{array} \right) \] 
\end{comment}
\end{proof}

\begin{comment}
Useful consequence: $z_{-\alpha}(X^2f(X), g(X)) \in P_\alpha(0)$. Proof: $z_{-\alpha}(X^2f(X), g(X)) = 
x_{-\beta}(-Xg) \cdot x_{-\alpha-\beta}(\epsilon X) \cdot x_{\alpha+\beta}(\epsilon X^3f^2g^2) \cdot x_{\alpha}(-X^2fg^2) \cdot 
z_{-\beta}(-Xg, -Xf) \cdot x_{\beta}(-X^3f^2g)  \cdot x_{-\alpha}(-X^2f) \cdot z_{-\alpha-\beta}(\epsilon X, -\epsilon Xfg)$,
where $\epsilon = N_{-\alpha-\beta,\beta}$. This is a product of generators of types: 1,1,2,2,1,1,4,1.
\end{comment}

@@

Let us verify the assertion for the generators of type (1) and (2). In the case $\beta - \alpha \in \Phi$ we obtain from Lemma~\ref{Zrels}:
\begin{align} z_{\beta}(Xf, g) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2fg) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f)     \cdot z_{\beta}(Xf, g), \label{eq3-1} \\
  z_{\beta}(f, Xg) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2fg ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf)     \cdot z_{\beta}(f, Xg), \label{eq3-2} \\
  z_{\alpha}(f, Xg) ^ {x_{-\alpha}(mX)} = & z_{\alpha}(f, X(g + m)). \end{align}
The expressions in the right hand side of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, g) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2fg ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2fg^2) \cdot z_{\beta}(Xf, g), \]
the latter expression, clearly, is a product of generators of type $4, 1, 1$.
\end{proof}

@@

\begin{df}
Define the map $S_\alpha(a, -) \colon P_\alpha(*) \to \St(\Phi, A[X, X^{-1}])$ on each coset $P_\alpha(m)$, $m \in M$ via the following formula:
\[ S_\alpha(a, g) = x_\alpha(aX^{-1})\cdot g \cdot x_\alpha\left(-\frac{aX^{-1}}{1 + am}\right) \cdot \{X, 1+ am\}_\alpha,\ g \in P_\alpha(m).\] 
\end{df}
It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the conjugation map $g \mapsto {}^{x_\alpha(aX^{-1})}\!g$.

\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator of $P_\alpha(0)$ lies in $P_\alpha(0)$. First of all, consider the case $\alpha + \beta \in \Phi$ (only possible for the generator of the first type):
\[ z_{\beta}(Xf, g) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- afg) \cdot x_{\beta+\alpha} (N_{\beta, \alpha}\cdot af)     \cdot z_{\beta}(Xf, g). \]
Clearly the expression in the right hand side is a product of elements of generators of type 2, 2, 1. 
Now consider the case $\alpha - \beta \in \Phi$ (possible for generators of type (1) and (2)).
\begin{align} z_{\beta}(Xf, g) ^ {x_{\alpha}(aX^{-1})} = &  x_{\alpha} (afg) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot afg^2) \cdot z_{\beta}(Xf, g), \\
z_{\beta}(f, Xg) ^ {x_{\alpha}(aX^{-1})} = & x_{\alpha} (afg) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot fXg^2a) \cdot z_{\beta}(f, Xg). \end{align}
The expressions in the right-hand side are products of generators of type 2, 2, 1 and 2, 1, 2.

Let us verify the assertion for the generator of type (4). Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$ and invoke~\eqref{Z5} of~\cref{Zrels} substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$:
\begin{multline} \nonumber
 z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(f a X^2) \cdot x_{-\alpha}(X^2f) \cdot \\ \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X),
\end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

It remains to verify the assertion for the remaining generator $z_\alpha(f, gX)$ of type 2. Substituting in~\eqref{Z5} of~\cref{Crels} $s = -\epsilon f$, $\xi = -\epsilon g$, $\eta=X$ and expressing $z_\alpha(f, gX)$ through other terms we obtain that
\begin{multline} \nonumber
 z_\alpha(f, gX) = 
 x_{\alpha+\beta}(\epsilon f X) \cdot
 x_{\beta}(-f gX^2) \cdot
 x_{-\beta}(f g) \cdot
 x_\alpha(f) \cdot \\ \cdot
 z_{\alpha+\beta}(-\epsilon fX, -\epsilon g) \cdot 
 z_{-\beta}(-fg, -X) \cdot
 x_{-\alpha-\beta}(-\epsilon f g^2 X) \cdot
 x_{-\alpha}(-fg^2X^2),
\end{multline}
where $\epsilon = N_{\alpha,\beta}$. 
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. 
\end{proof}

\begin{lemma} \label{lem:Tulenbayev-formula}
 $x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}$.
\end{lemma}
\begin{proof}
Since $\Phi$ is nonsymplectic, we can choose some $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$.
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}. \qedhere
\end{multline*}
\end{proof}

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$. \end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right).\qedhere \end{multline}
\end{proof}

>>>>

{\bf Warning}: This definition is somewhat incomplete, $P_\alpha(0)$ should also contain certain symbols, e.\,g. those from~\cref{lem:Tulenbayev-formula} below. 
I'll fix this later.

@@

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of type 3. \end{proof}

The above lemma allows us to give the following alternative characterization for the subgroup $P_\alpha(*)$:
 it consists precisely of those elements $g \in \St(\Phi, A[X], M[X])$ for which $g(0)$ lies in the subgroup generated by
 elements $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$ 
 and $Z_+(\alpha)$, $Z_0(\alpha)$ are as in~\cite[Proposition~5.7]{St71}.

@@

Let us verify the assertion for the generators of type (1) and (2). In the case $\beta - \alpha \in \Phi$ we obtain from Lemma~\ref{Zrels}:
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi), \label{eq3-1} \\
  z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi), \label{eq3-2} \\
  z_{\alpha}(f, X\xi) ^ {x_{-\alpha}(mX)} = & z_{\alpha}(f, X(\xi + m)). \end{align}
The expressions in the right hand side of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2f\xi ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2f\xi^2) \cdot z_{\beta}(Xf, \xi), \]
the latter expression, clearly, is a product of generators of type $4, 1, 1$.
\end{proof}

@@

\begin{df} Define the map $S_\alpha(a, -) \colon P_\alpha(*) \to \St(\Phi, A[X, X^{-1}])$ on each coset $P_\alpha(m)$, $m \in M$ via the following formula:
\[ S_\alpha(a, g) = x_\alpha(aX^{-1})\cdot g \cdot x_\alpha\left(-\frac{aX^{-1}}{1 + am}\right) \cdot \{X, 1+ am\}_\alpha,\ g \in P_\alpha(m).\] \end{df}
Right from the definition of $S_\alpha(a, -)$ it is easy to verify that for $g_1 \in P_\alpha(m)$, $g_2 \in P_\alpha(*)$ holds
\[ S_\alpha(a, g_1\cdot g_2) = S_\alpha(a, g_1) \cdot S_\alpha\left(\frac{a}{1+am}, g_2\right).\]

It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the conjugation map $g \mapsto {}^{x_\alpha(aX^{-1})}\!g$.

\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator of $P_\alpha(0)$ lies in $P_\alpha(0)$. 
First of all, consider the case $\alpha + \beta \in \Phi$ (only possible for the generator of the first type):
\[ z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- afg) \cdot x_{\beta+\alpha} (N_{\beta, \alpha}\cdot af)     \cdot z_{\beta}(Xf, \xi). \]
Clearly the expression in the right hand side is a product of elements of generators of type 2, 2, 1. 
Now consider the case $\alpha - \beta \in \Phi$ (possible for generators of type (1) and (2)).
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = &  x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot af\xi^2) \cdot z_{\beta}(Xf, \xi), \\
z_{\beta}(f, X\xi) ^ {x_{\alpha}(aX^{-1})} = & x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot Xf\xi^2a) \cdot z_{\beta}(f, X\xi). \end{align}
The expressions in the right-hand side are products of generators of type 2, 2, 1 and 2, 1, 2.

Let us verify the assertion for the generator of type (4).
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$ and invoke~\eqref{Z5} of~\cref{Zrels} substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$:
\begin{multline} \nonumber z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(X^2 f a) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

It remains to verify the assertion for the remaining generator $z_\alpha(f, gX)$ of type 2.
Substituting in~\eqref{Z5} of~\cref{Crels} $s := -\epsilon f$, $\xi = -\epsilon g$, $\eta=X$ we obtain that
\begin{multline} \nonumber z_{\alpha+\beta}(-\epsilon fX, -\epsilon \xi) = x_\alpha( -f) x_{-\beta}(- f \xi) \cdot x_{\beta}( X^2f \xi) \cdot x_{\alpha+\beta}(-X\epsilon f) \cdot \\
 \cdot z_\alpha(f,  X\xi) \cdot x_{-\alpha}(X^2f\xi^2) \cdot x_{-\alpha-\beta}(\epsilon Xf \xi^2) \cdot z_{-\beta}(f\xi, -X), \end{multline}
where $\epsilon = N_{\alpha,\beta}$. 
We can express $z_{\alpha}(f, X\xi)$ through other terms and obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2) \cdot \end{multline}
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. \end{proof}

\begin{lemma} \label{lem:Tulenbayev-formula} One has
 \[x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot 
 x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}.\] \end{lemma}
\begin{proof} Since $\Phi$ is nonsymplectic, we can choose some $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$.
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}. \qedhere
\end{multline*}
\end{proof}

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$.
Consequently one has \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right). \end{multline}
The second assertion can be verified in a straightforward manner. \end{proof}
